Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}

library(ggplot2)
pf = read.csv('E:\\Study\\coursera\\Udacity\\ud651DataAnalysisWithR\\pseudo_facebook.tsv', sep = '\t')

qplot(x = age, y = friend_count, data = pf)
qplot(age, friend_count, data = pf)
```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}

ggplot(data = pf, aes(x = age, y = friend_count)) + geom_point()
summary(pf$age)

ggplot(data = pf, aes(x = age, y = friend_count)) + geom_point() + xlim(13,90)



```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(data = pf, aes(x = age, y = friend_count)) + geom_point(alpha = 1/20) + xlim(13,90)

ggplot(data = pf, aes(x = age, y = friend_count)) + geom_jitter(alpha = 1/20) + xlim(13,90)

```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}

ggplot(data = pf, aes(x = age, y = friend_count)) + geom_point(alpha = 1/20) + xlim(13,90) + coord_trans(y = 'sqrt')
ggplot(data = pf, aes(x = age, y = friend_count)) + geom_point(alpha = 1/20, position = position_jitter(h = 0)) + xlim(13,90) + coord_trans(y = 'sqrt')

```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}

```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}

ggplot(data = pf, aes(x = age, y = friendships_initiated)) + geom_point(alpha = 1/20) + xlim(13,90)
ggplot(data = pf, aes(x = age, y = friendships_initiated)) + geom_point(alpha = 1/20, position = 'jitter') + xlim(13,90)
ggplot(data = pf, aes(x = age, y = friendships_initiated)) + geom_point(alpha = 1/20, position = position_jitter(h = 0)) + xlim(13,90)
ggplot(data = pf, aes(x = age, y = friendships_initiated)) + geom_point(alpha = 1/20, position = position_jitter(h = 0)) + xlim(13,90) + coord_trans(y = 'sqrt')

```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}

#install.packages('dplyr')
library('dplyr')

#filter()
#group_by()
#mutate()
#arrange()

age_groups = group_by(pf,age)
pf.fc_by_age = summarise(age_groups, 
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n())
pf.fc_by_age = arrange(pf.fc_by_age, age)
head(pf.fc_by_age)

```

## Conditional Means Alternate Code
```{r}

pf.fc_by_age = pf %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age)

head(pf.fc_by_age)

```

Create your plot!

```{r Conditional Means Plot}

ggplot(data = pf.fc_by_age, aes(x = age, y = friend_count_mean)) + geom_point() 

ggplot(data = pf.fc_by_age, aes(x = age, y = friend_count_mean)) + geom_line() 


```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}

ggplot(data = pf, aes(x = age, y = friendships_initiated)) + geom_point(alpha = 1/20, position = position_jitter(h = 0), color = 'orange') + xlim(13,90) + coord_trans(y = 'sqrt')

ggplot(data = pf, aes(x = age, y = friendships_initiated)) + geom_point(alpha = 1/20, position = position_jitter(h = 0), color = 'orange') + xlim(13,90) + coord_trans(y = 'sqrt') + geom_line(stat = 'summary', fun.y = mean)

ggplot(data = pf, aes(x = age, y = friendships_initiated)) + geom_point(alpha = 1/20, position = position_jitter(h = 0), color = 'orange') + xlim(13,90) + coord_trans(y = 'sqrt') + geom_line(stat = 'summary', fun.y = mean) + geom_line(stat = 'summary', fun.y = quantile, fun.args = (probs = .1), linetype = 2, color = 'blue' )

ggplot(data = pf, aes(x = age, y = friendships_initiated)) + geom_point(alpha = 1/20, position = position_jitter(h = 0), color = 'orange') + xlim(13,90) + coord_trans(y = 'sqrt') + geom_line(stat = 'summary', fun.y = mean) + geom_line(stat = 'summary', fun.y = quantile, fun.args = (probs = .1), linetype = 2, color = 'blue' ) + geom_line(stat = 'summary', fun.y = quantile, fun.args = (probs = .9), linetype = 2, color = 'blue' )

ggplot(data = pf, aes(x = age, y = friendships_initiated)) + geom_point(alpha = 1/20, position = position_jitter(h = 0), color = 'orange') + xlim(13,90) + coord_trans(y = 'sqrt') + geom_line(stat = 'summary', fun.y = mean) + geom_line(stat = 'summary', fun.y = quantile, fun.args = (probs = .1), linetype = 2, color = 'blue' ) + geom_line(stat = 'summary', fun.y = quantile, fun.args = (probs = .9), linetype = 2, color = 'blue' ) + geom_line(stat = 'summary', fun.y = median,  color = 'blue' )

ggplot(data = pf, aes(x = age, y = friendships_initiated)) + geom_point(alpha = 1/20, position = position_jitter(h = 0), color = 'orange') + xlim(13,90) + coord_trans(y = 'sqrt') + geom_line(stat = 'summary', fun.y = mean) + geom_line(stat = 'summary', fun.y = quantile, fun.args = (probs = .1), linetype = 2, color = 'blue' ) + geom_line(stat = 'summary', fun.y = quantile, fun.args = (probs = .9), linetype = 2, color = 'blue' ) + geom_line(stat = 'summary', fun.y = median,  color = 'blue' ) + coord_cartesian(xlim = c(13,40))

ggplot(data = pf, aes(x = age, y = friendships_initiated)) + geom_point(alpha = 1/20, position = position_jitter(h = 0), color = 'orange') + geom_line(stat = 'summary', fun.y = mean) + geom_line(stat = 'summary', fun.y = quantile, fun.args = (probs = .1), linetype = 2, color = 'blue' ) + geom_line(stat = 'summary', fun.y = quantile, fun.args = (probs = .9), linetype = 2, color = 'blue' ) + geom_line(stat = 'summary', fun.y = median,  color = 'blue' ) + coord_cartesian(xlim = c(13,70), ylim = c(0,1000))

```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}

cor.test(pf$age, pf$friend_count, method = 'pearson')
with(pf, cor.test(age, friend_count, method = 'pearson'))

```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age <= 70), cor.test(age, friend_count))

with(subset(pf, age <= 70), cor.test(age, friend_count, method = 'pearson'))
```

***

### Correlation Methods
```{r Correlation Methods}

with(subset(pf, age <= 70), cor.test(age, friend_count, method = 'spearman'))
```

***

## Create Scatterplots
Notes:

```{r}
#ggplot(data = pf, aes(x = age, y = friendships_initiated)) + geom_point(alpha = 1/20, position = position_jitter(h = 0), color = 'orange') + geom_line(stat = 'summary', fun.y = mean) + geom_line(stat = 'summary', fun.y = quantile, fun.args = (probs = .1), linetype = 2, color = 'blue' ) + geom_line(stat = 'summary', fun.y = quantile, fun.args = (probs = .9), linetype = 2, color = 'blue' ) + geom_line(stat = 'summary', fun.y = median,  color = 'blue' ) + coord_cartesian(xlim = c(13,70), ylim = c(0,1000))


ggplot(data = pf, aes(x = www_likes_received, y = likes_received)) + geom_point()

ggplot(data = pf, aes(x = www_likes_received, y = likes_received)) + geom_point() + xlim(0,quantile(pf$www_likes_received,0.95)) + ylim(0,quantile(pf$likes_received,0.95)) + geom_smooth(method = 'lm', color = 'red')


```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
with(pf, cor.test(www_likes_received, likes_received))

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}

#install.packages('alr3')
library(alr3)
data("Mitchell")
?Mitchell
```

Create your plot!

```{r Temp vs Month}

ggplot(data = Mitchell, aes(x = Month, y = Temp)) + geom_point()

```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Temp, Mitchell$Month)
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(data = Mitchell, aes(x = Month, y = Temp)) + geom_point() + scale_x_continuous(breaks = 12*(1:17))

ggplot(data = Mitchell, aes(x = Month, y = Temp)) + geom_point() + scale_x_continuous(breaks =seq(0,203,12))

```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
#ggplot(data = pf, aes(x = age, y = friendships_initiated)) + geom_point(alpha = 1/20, position = position_jitter(h = 0), color = 'orange') + geom_line(stat = 'summary', fun.y = mean) + geom_line(stat = 'summary', fun.y = quantile, fun.args = (probs = .1), linetype = 2, color = 'blue' ) + geom_line(stat = 'summary', fun.y = quantile, fun.args = (probs = .9), linetype = 2, color = 'blue' ) + geom_line(stat = 'summary', fun.y = median,  color = 'blue' ) + coord_cartesian(xlim = c(13,70), ylim = c(0,1000))
# Create a new variable, 'age_with_months', in the 'pf' data frame.
# Be sure to save the variable in the data frame rather than creating
# a separate, stand-alone variable. You will need to use the variables
# 'age' and 'dob_month' to create the variable 'age_with_months'.

# Assume the reference date for calculating age is December 31, 2013.
pf$age_with_months = (12 - pf$dob_month)/12 + pf$age

ggplot(data = pf, aes(x = age_with_months, y = friendships_initiated)) + geom_point(alpha = 1/20, position = position_jitter(h = 0), color = 'orange') + geom_line(stat = 'summary', fun.y = mean) + geom_line(stat = 'summary', fun.y = quantile, fun.args = (probs = .1), linetype = 2, color = 'blue' ) + geom_line(stat = 'summary', fun.y = quantile, fun.args = (probs = .9), linetype = 2, color = 'blue' ) + geom_line(stat = 'summary', fun.y = median,  color = 'blue' )

age_with_month_groups = group_by(pf,age_with_months)
pf.fc_by_age_months = summarise(age_with_month_groups, 
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n())
pf.fc_by_age_months = arrange(pf.fc_by_age_months)

ggplot(data = pf.fc_by_age_months, aes(x=age_with_months, y = friend_count_mean)) +geom_point()

```

***

### Age with Months Means

```{r Age with Months Means}
ggplot(data = pf.fc_by_age_months, aes(x=age_with_months, y = friend_count_mean)) +geom_line() + coord_cartesian(xlim = c(13,70))


```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
ggplot(data = subset(pf.fc_by_age_months, age_with_months <71), aes(x=age_with_months, y = friend_count_mean)) +geom_line()
ggplot(data = subset(pf.fc_by_age, age <71), aes(x=age, y = friend_count_mean)) +geom_line()

p1 = ggplot(data = subset(pf.fc_by_age_months, age_with_months <71), aes(x=age_with_months, y = friend_count_mean)) +geom_line()
p2 = ggplot(data = subset(pf.fc_by_age, age <71), aes(x=age, y = friend_count_mean)) +geom_line()
library(gridExtra)
grid.arrange(p2,p1,ncol=1)

p1 = ggplot(data = subset(pf.fc_by_age_months, age_with_months <71), aes(x=age_with_months, y = friend_count_mean)) +geom_line()
p2 = ggplot(data = subset(pf.fc_by_age, age <71), aes(x=age, y = friend_count_mean)) +geom_line()
p3 = ggplot(data = subset(pf, age <71), aes(x=round(age/5)*5, y = friend_count)) +geom_line(stat = 'summary', fun.y = mean)
library(gridExtra)
grid.arrange(p1,p2,p3,ncol=1)


p1 = ggplot(data = subset(pf.fc_by_age_months, age_with_months <71), aes(x=age_with_months, y = friend_count_mean)) +geom_line() + geom_smooth()
p2 = ggplot(data = subset(pf.fc_by_age, age <71), aes(x=age, y = friend_count_mean)) +geom_line() + geom_smooth()
p3 = ggplot(data = subset(pf, age <71), aes(x=round(age/5)*5, y = friend_count)) +geom_line(stat = 'summary', fun.y = mean)
library(gridExtra)
grid.arrange(p1,p2,p3,ncol=1)
```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

