Lesson 5
========================================================
### Scatterplots
Notes:

```{r Scatterplots}

library(ggplot2)
pf = read.csv('E:\\Study\\coursera\\Udacity\\ud651DataAnalysisWithR\\pseudo_facebook.tsv', sep = '\t')

qplot(x = age, y = friend_count, data = pf)
```

***
### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}

ggplot(aes(x = gender, y = age), data = subset(pf, !is.na(gender))) + geom_boxplot() 
#add average
ggplot(aes(x = gender, y = age), data = subset(pf, !is.na(gender))) + geom_boxplot() + stat_summary(fun.y = mean, geom = 'point', shape = 4)

ggplot(aes(x = age, y = friend_count), data = subset(pf, !is.na(gender))) + geom_line(aes(color = gender), stat = 'summary', fun.y = median)

# Write code to create a new data frame,
# called 'pf.fc_by_age_gender', that contains
# information on each age AND gender group.
# The data frame should contain the following variables:

#    mean_friend_count,
#    median_friend_count,
#    n (the number of users in each age and gender grouping)
library(dplyr)

pf.fc_by_age = pf %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age)

#chain functions together
pf.fc_by_age_gender = subset(pf, !is.na(pf$gender)) %>%
  group_by(age,gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %>%
  ungroup() %>%
  arrange(age)
head(pf.fc_by_age_gender)
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot(data = pf.fc_by_age_gender, aes(x = age, y = median_friend_count)) +geom_line(aes(color = gender))

```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
#install.packages('reshape2')
library(reshape2)

pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender, age ~ gender, value.var = 'median_friend_count')


```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(data = pf.fc_by_age_gender.wide, aes(x = age, y = female/male)) + geom_line() + geom_hline(yintercept = 1, alpha = 0.3, linetype = 2)
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}

pf$year_joined <- floor(2014 - pf$tenure/365)
table(pf$year_joined)
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
pf$year_joined.bucket <- cut(pf$year_joined, c(2004, 2009, 2011, 2012,2014))
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
ggplot(data = subset(pf, !is.na(pf$year_joined.bucket)), aes(x = age, y = friend_count)) +geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(data = subset(pf, !is.na(pf$year_joined.bucket)), aes(x = age, y = friend_count)) +geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean)+geom_line( stat = 'summary', fun.y = mean, linetype = 3)
```

***

### Friending Rate
Notes:

```{r Friending Rate}
pf$friend_rate = pf$friend_count/pf$tenure
summary(subset(pf, !is.infinite(pf$friend_rate ))$friend_rate)

#althernative code
with(subset(pf, tenure >= 1), summary(friend_count / tenure))
```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}
ggplot(data = subset(pf, tenure >= 1), aes(x = tenure, y = friendships_initiated / tenure)) +geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean)+geom_line( stat = 'summary', fun.y = mean, linetype = 3)
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

# Instead of geom_line(), use geom_smooth() to add a smoother to the plot.
# You can use the defaults for geom_smooth() but do color the line
# by year_joined.bucket
ggplot(data = subset(pf, tenure >= 1), aes(x = tenure, y = friendships_initiated / tenure)) +geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) + geom_smooth(linetype = 3)
#below is the right answer
ggplot(data = subset(pf, tenure >= 1), aes(x = tenure, y = friendships_initiated / tenure)) +geom_smooth(aes(color = year_joined.bucket))

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo <- read.csv('E:\\Study\\coursera\\Udacity\\ud651DataAnalysisWithR\\yogurt.csv')
str(yo)

#change the id from an int to a factor
yo$id <- factor(yo$id)
str(yo)

ggplot(data = yo, aes(x=price)) + geom_histogram()

#alternative
qplot(data = yo, x = price, fill = I('#F79420'))

qplot(data = yo, x = price, fill = I('#F79420'), binwidth = 10)
```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
summary(yo)
length(unique(yo$price))
table(yo$price)

# Create a new variable called all.purchases,
# which gives the total counts of yogurt for
# each observation or household.
yo$all.purchases = yo$strawberry + yo$blueberry + yo$pina.colada + yo$plain + yo$mixed.berry

#alternative
yo <- transform(yo, all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)

qplot(x = all.purchases, data = yo, binwidth = 1, fill = I('#099DD9'))
```

***

### Prices over Time
Notes:

```{r Prices over Time}
ggplot(data = yo, aes(x = time, y = price)) + geom_point()

#in the video
ggplot(data = yo, aes(x = time, y = price)) + geom_jitter(alpha = 1/4, shape = 21, fill = I('#F79420'))


```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
#set.seed function to make this result reproducible.
set.seed(423)
sample.ids <- sample(levels(yo$id), 25)

ggplot(aes(x = time, y = price), data = subset(yo, id %in% sample.ids)) + facet_wrap(~ id) + geom_line() + geom_point(aes(size = all.purchases), pch = 1)

#id %in% sample.ids, loop the id
```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***

```{r scatterplot matrix}
#install.packages('GGally')
library(GGally)
theme_set(theme_minimal(20))

#set the seed for reproducible results
set.seed(1836)
pf_subset <- pf[,c(2:15)] #do not use user id and year_joined.bucket
names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000),])

```

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("E:\\Study\\coursera\\Udacity\\ud651DataAnalysisWithR\\nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

